\LARGE \bf Planning with Attitude

### 2020-10-10 paper/main.tex
Minor quaternion notation problems

    Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
    rotation from the body to inertial frames ${}^N q^N$, expressed as a quaternion. We
    can define Wahba's loss function as the following: 
    \begin{align} \label{eq:wahba_loss}
        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
    \end{align} 
    where $r_i(q)$ is the residual vector.

@@

    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$,
    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
    \eqref{eq:quat_jacobian}:
    \begin{equation} 
        \bar{J} = \pdv{r(q \otimes \varphi{\phi})}{\phi} = \pdv{r(q)}{q} G(q).
    \end{equation}
    We obtain our step---this time in the actual tangent space---as before:
    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
    step using the correct notion of composition for the group: 
    $q_{k+1} = q_k \otimes \varphi{\phi}$. This ``multiplicative'' Newton algorithm is 
    summarized in Algorithm \ref{alg:mgn}.

    \begin{algorithm} 
    	\begin{algorithmic}[1]
    		\caption{Multiplicative Gauss-Newton Method}\label{alg:mgn}
    		\State $k = 0$
    		\While{significant progress}
    		    \State $\bar{J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
    		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
    		    \State $q_{k+1} = q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
    		    \State $k = k + 1$
    		\EndWhile
    	\end{algorithmic}
    \end{algorithm}

@@

        \subsubsection{Error Quadratic}
            Rather than simple subtraction, we can use a quadratic function on the
            three-parameter error state \eqref{eq:state_error}:
            \begin{equation} \label{eq:error_quadratic}
                J_\text{err} = \half \phi^T Q \phi 
                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
                        \left(\varphi^{-1}(\delta q)\right).
            \end{equation}
            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
            and Hessian of \eqref{eq:error_quadratic} are
            \begin{equation}
                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
            \end{equation}
            \begin{multline}
                \nabla^2 J_\text{err} = 
                    G(\delta q)^T \! \left(
                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
                    + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
            \end{multline}
            where, for the Cayley map,
            \begin{equation}
                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
                    q_v \;\; & -\frac{1}{q_s} I_3
                \end{bmatrix}
            \end{equation}
            \begin{equation}
                \nabla D = \pdv{q}(D(q)^T Q \phi) 
                = -\frac{1}{q_s^2} \begin{bmatrix} 
                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
                                   Q \phi & 0 \\
                \end{bmatrix}.
            \end{equation}

@@

    \subsection{Satellite Attitude Keep-Out}
    A spacecraft with flexible appendages modes must perform a 150 degree slew, while making sure that a startracker does not look within 40 degrees of the sun. The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}. We can guarantee the startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
    \begin{equation}
        \langle ^N A ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
    \end{equation}
    From this constraint, an eigen-axis slew trajectory planner may violate this constraint \cite{markley2014fundamentals}. ALTRO is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker line of sight constraint.
        \begin{figure}
            \centering
            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
            \caption{Visualization of the flexible spacecraft slew with a dedicated keep out zone. Attitude is parameterized with a Rodrigues parameter to visualize in 3D. The constraint surface represents attitudes where the startracker is within 40$^\circ$ of the sun. The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}
        \end{figure}

@@

>>>>

    Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
    rotation from the body to inertial frames ${}^N q^N$, expressed as a quaternion. We
    can define Wahba's loss function as the following: 
    \begin{align} \label{eq:wahba_loss}
        L &= \sum_i w_i \norm{{}^N v_i - {}^N A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
    \end{align} 
    where $r_i(q)$ is the residual vector.

@@

    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$,
    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
    \eqref{eq:quat_jacobian}:
    \begin{equation} 
        \bar{J} = \pdv{r(q \otimes \varphi{\phi})}{\phi} = \pdv{r(q)}{q} G(q).
    \end{equation}
    We obtain our step---this time in the actual tangent space---as before:
    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
    step using the correct notion of composition for the group: 
    $q_{k+1} = L(q_k) \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
    summarized in Algorithm \ref{alg:mgn}.

    \begin{algorithm} 
    	\begin{algorithmic}[1]
    		\caption{Multiplicative Gauss-Newton Method}\label{alg:mgn}
    		\State $k = 0$
    		\While{significant progress}
    		    \State $\bar{J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
    		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
    		    \State $q_{k+1} = L(q_k) \varphi(\phi)$ \Comment{apply step multiplicatively}
    		    \State $k = k + 1$
    		\EndWhile
    	\end{algorithmic}
    \end{algorithm}

@@

        \subsubsection{Error Quadratic} \label{sec:error_quadratic}
            Rather than simple subtraction, we can use a quadratic function on the
            three-parameter error state \eqref{eq:state_error}:
            \begin{equation} \label{eq:error_quadratic}
                J_\text{err} = \half \phi^T Q \phi 
                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
                        \left(\varphi^{-1}(\delta q)\right).
            \end{equation}
            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
            and Hessian of \eqref{eq:error_quadratic} are
            \begin{equation}
                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
            \end{equation}
            \begin{multline}
                \nabla^2 J_\text{err} = 
                    G(\delta q)^T \! \left(
                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
                    + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
            \end{multline}
            where, for the Cayley map,
            \begin{equation}
                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
                    q_v \;\; & -\frac{1}{q_s} I_3
                \end{bmatrix}
            \end{equation}
            \begin{equation}
                \nabla D = \pdv{q}(D(q)^T Q \phi) 
                = -\frac{1}{q_s^2} \begin{bmatrix} 
                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
                                   Q \phi & 0 \\
                \end{bmatrix}.
            \end{equation}

@@

    \subsection{Satellite Attitude Keep-Out}
    A spacecraft with flexible appendages must perform a 150 degree slew, while
    making sure that a startracker does not look within 40 degrees of the sun. The
    flexible body spacecraft dynamics are based on the truncated hybrid coordinate model,
    with the three largest modes captured \cite{Tracy2020}. We can guarantee the
    startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the
    inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
    \begin{equation}
        \langle ^N A ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
    \end{equation}
    From this constraint, an eigen-axis slew trajectory planner may violate this
    constraint \cite{markley2014fundamentals}. ALTRO is able to reason about the
    nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker
    line of sight constraint. The resulting attitude trajectory is shown Figure
    \ref{{fig:satellite}}
        \begin{figure}
            \centering
            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
            \caption{Visualization of the flexible spacecraft slew with a dedicated keep out zone. Attitude is parameterized with a Rodrigues parameter to visualize in 3D. The constraint surface represents attitudes where the startracker is within 40$^\circ$ of the sun. The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}
            \label{fig:satellite}
        \end{figure}

@@

        We also compared using a quadratic cost on the 3-dimensional quaternion error
        described in Section \ref{sec:error_quadratic} to the geodesic cost function in
        Section \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of 
        iterations and computation time, taking only 41 iterations and
